Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
done scalability

\emph{Scalability} corresponds to the ability of the architecture to
cope with high workloads. It is measured by counting the potential
scalability issues in the architectures, i.e. the missing scalability
features. Features are evaluated using a 3-level metric: 0 means that
the feature is very easy to enable, 1 means that it can be implemented
but with some difficulty, and 2 means that the feature cannot
realistically be implemented. \closednote{Naj}{how do you metricize
  "nonesensical amount fo effors?"}{Tristan}{You don't :) Rephrased to
  'cannot realistically be implemented'. At some point it's a matter
  of common sense and experience to draw the line between what is
  easily implemented and what is more tricky. I would like to avoid
  'impossible' because nothing is impossible to implement if enough
  effort is put.} Four different features are identified:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Multiple engine instances (\texttt{S$_1$}): possibility to have
  more than 1 engine instance in the architecture. Workflow engines
  may require important amounts of resources when several workflows,
  or large workflows, are executed. At some point, it may be required
  for the science gateway to distribute the load among several
  engines. \texttt{S$_1$}=0 when adding a new engine instance is a
  fully automated process, i.e. the workflow engine only has to be
  started. In this case, elastic engines are possible, i.e. some kind
  of auto-scaling mechanism can be implemented to control the number
  of engine instances in the architecture. \texttt{S$_1$}=1 when
  adding a new engine instance requires some form of manual
  intervention (e.g., two-factor authentication), which prevents easy
  implementations of elastic engines. \texttt{S$_1$}=2 when new engine
  instances cannot be added.
\item Distributed engines (\texttt{S$_2$}): possibility to distribute
  the execution \emph{of a single workflow} among different engine
  instances. In our scope, this feature focuses on the capabilities of
  the architecture rather than these of the workflow
  engine. \texttt{S$_2$}=0 when distributed engines are possible in
  the architecture, \texttt{S$_2$}=1 when they require specific
  developments in the workflow engine.  \closednote{Rafael}{$S_2$
    would be 2 if the paradigm does not allow to distribute the
    orchestration? e.g., workflow executions implemented within MPI
    jobs, in Pegasus PMC and dispel4Py.}{Tristan}{The problem is that
    it is very difficult to define ``does not allow to''. For
    instance, why couldn't you hack Pegasus PMC so much that it
    becomes possible to distribute workflow executions in several MPI
    jobs coordinated by an MPI job? Moreover, S2=2 is not used so I'd
    prefer to leave it like this, unless you have a suggestion to
    reflect this in the text.}
\item Task scheduling (\texttt{S$_3$}): task scheduling is a difficult
  issue that depends more on the implementation of specific algorithms
  in the science gateway, workflow engine and infrastructure than on
  the architecture used to integrate the workflow engine in the
  science gateway. Some architectures, however, complicate the task
  scheduling problem by introducing additional software layers or
  creating tasks with specific characteristics. \texttt{S$_3$}=0 when
  the architecture does not introduce any additional complexity to the
  scheduling problem, and \texttt{S$_3$}=1 otherwise.
\end{itemize}
The 3 metrics are summed to obtain a global measure of the scalability
issues that are present in the architecture.

>>>>

\emph{Scalability} corresponds to the ability of an architecture to
cope with high workloads through the following mechanisms: starting multiple engine instances, distributing engines and task scheduling. 
This is measured by assessing the potential
scalability difficulties due to (partial) absence of a given feature in the architecture.  
Features are evaluated using a 3-level metric: 0 means that
the feature is very easy to enable, 1 means that it can be implemented
but with some difficulty, and 2 means that the feature cannot
 be implemented in a realistic scenario. \closednote{Naj}{how do you metricize
  "nonesensical amount fo effors?"}{Tristan}{You don't :) Rephrased to
  'cannot realistically be implemented'. At some point it's a matter
  of common sense and experience to draw the line between what is
  easily implemented and what is more tricky. I would like to avoid
  'impossible' because nothing is impossible to implement if enough
  effort is put.} Three different features are identified:
\item Multiple engine instances (\texttt{S$_1$}): measures the possibility \note{sil}{are we measuring possibility or potential? i think the second, because it is hypothetical. possibility brings me the idea of it is possible or not - which would be zero to almost all these features} of having
  more than one engine instance in the architecture simultaneously. Workflow engines
  may require important amounts of resources when several workflows,
  or large workflows, are executed. At some point, it may be required
  for the science gateway to means that adding a new engine instance is a
  fully automated process, i.e. the workflow engine only has to be
  started. In this case, elastic engines are possible, i.e. some kind
  of auto-scaling mechanism can be implemented to control the number
  of engine instances in the architecture. \texttt{S$_1$}=1 means that
  adding a new engine instance requires some form of manual
  intervention (e.g., two-factor authentication \note{sil}{what is this?}), which prevents easy
  implementations of elastic engines. \texttt{S$_1$}=2 means that a new engine
  instances cannot be added.
\item Distributed engines (\texttt{S$_2$}): measures the possibility of distributing
  the execution \emph{of a single workflow} among different engine
  instances. In our scope, this feature focuses on the capabilities of
  the architecture rather than these of the workflow
  engine. \texttt{S$_2$}=0 means that distributed engines are possible in
  the architecture, and \texttt{S$_2$}=1 that they require specific
  developments in the workflow engine.  \closednote{Rafael}{$S_2$
    would be 2 if the paradigm does not allow to distribute the
    orchestration? e.g., workflow executions implemented within MPI
    jobs, in Pegasus PMC and dispel4Py.}{Tristan}{The problem is that
    it is very difficult to define ``does not allow to''. For
    instance, why couldn't you hack Pegasus PMC so much that it
    becomes possible to distribute workflow executions in several MPI
    jobs coordinated by an MPI job? Moreover, S2=2 is not used so I'd
    prefer to leave it like this, unless you have a suggestion to
    reflect this in the text.}
\item Task scheduling (\texttt{S$_3$}): measures ??? \note{sil}{need to be more specific about what is being called task scheduling here. also explain why it is difficult, or dont say anything if it is common knowledge}. Task scheduling typically depends more on the implementation of specific algorithms
  in the science gateway, workflow engine and infrastructure than on
  the architecture used to integrate the workflow engine in the
  science gateway. Some architectures, however, complicate the task
  scheduling problem by introducing additional software layers or
  creating tasks with specific characteristics. \texttt{S$_3$}=0 means that
  the architecture does not introduce any additional complexity to the
  scheduling problem, and \texttt{S$_3$}=1 otherwise.
These three metrics are summed to obtain a global measure of the scalability
potential of the architecture.
